Add Plant Functional Type (PFT) labels

### 2021-04-23 R/add_plant_functional_type.Rmd
making progress, using adam's tips about group to identify needle vs broad leaf

```{r manually_add_pft_labels}
# list unique taxon labels, add an empty columns to fill with plant functional type (PFT) and common name labels.
taxon_pft_lookup <- data.frame(genusSpecies = unique(cleaned_spectra$genusSpecies),
                               commonName = NA,
                               pft_MRA = NA,
                               pft_MODIS = NA,
                               phenology = NA,
                               leafType = NA,
                               growthForm = NA) %>% 
  # arrange alphabetically
  dplyr::arrange(genusSpecies)

@@

  else if(genusSpecies == "Acacia greggii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub" 
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "catclaw acacia"}

@@

  else if(genusSpecies == "Aloysia wrightii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "Wright's beebrush"}

@@

  else if(genusSpecies == "Artemisia tridentata"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "big sagebrush"}

  else if(genusSpecies == "Atriplex canescens"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "fourwing saltbush"}

  else if(genusSpecies == "Atriplex confertifolia"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "shadscale saltbush"}

@@

  else if(genusSpecies == "Bourreria sp."){
    taxon_pft_lookup$pft_MRA[i] = "NA"  # taxon class is too broad to specify 
    taxon_pft_lookup$pft_MODIS[i] = "NA"
    taxon_pft_lookup$phenology[i] = "NA"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "strongbark"}

@@

  else if(genusSpecies == "Chrysothamnus viscidiflorus"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "yellow rabbitbrush"}

@@

  else if(genusSpecies == "Cornus drummondii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub" 
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "roughleaf dogwood"}

@@

  else if(genusSpecies == "Corylus cornuta"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub" 
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "beaked hazelnut"}

@@

  else if(genusSpecies == "Ephedra trifurca"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "longleaf jointfir"}

  else if(genusSpecies == "Ericameria laricifolia"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "turpentine bush"}

@@

  else if(genusSpecies == "Gaultheria shallon"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "salal"}

@@

  else if(genusSpecies == "Gutierrezia sarothrae"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "NA"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "broom snakeweed"}

@@

  else if(genusSpecies == "Juniperus osteosperma"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "evergreen_needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf" # scale-like needle leaves
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "Utah juniper"}

  else if(genusSpecies == "Juniperus virginiana"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "evergreen_needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf" # scale-like needle leaves
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "red cedar"}

  else if(genusSpecies == "Larix sp."){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous" # larch is an unusual genus with deciduous conifers
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "larch"}

  else if(genusSpecies == "Larrea tridentata"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = ""
    taxon_pft_lookup$phenology[i] = ""
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = ""
    taxon_pft_lookup$commonName[i] = "creosote bush"}

  else if(genusSpecies == "Liquidambar styraciflua"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = ""
    taxon_pft_lookup$phenology[i] = ""
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = ""
    taxon_pft_lookup$commonName[i] = "sweetgum"}

  else if(genusSpecies == "Liriodendron tulipifera"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = ""
    taxon_pft_lookup$phenology[i] = ""
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = ""
    taxon_pft_lookup$commonName[i] = "tuliptree"}

  else if(genusSpecies == "Lonicera maackii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = ""
    taxon_pft_lookup$phenology[i] = ""
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = ""
    taxon_pft_lookup$commonName[i] = "Amur honeysuckle"}

  else if(genusSpecies == "Maclura pomifera"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"   # small deciduous tree or large shrub
    taxon_pft_lookup$pft_MODIS[i] = ""
    taxon_pft_lookup$phenology[i] = ""
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = ""
    taxon_pft_lookup$commonName[i] = "Osage-orange"}

 else if(genusSpecies == "Magnolia sp."){
    taxon_pft_lookup$pft_MRA[i] = "NA"
    taxon_pft_lookup$pft_MODIS[i] = ""
    taxon_pft_lookup$phenology[i] = ""
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = ""
    taxon_pft_lookup$commonName[i] = "magnolia"}

  else if(genusSpecies == "Mahonia nervosa"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"      # evergreen shrub
    taxon_pft_lookup$pft_MODIS[i] = ""
    taxon_pft_lookup$phenology[i] = ""
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = ""
    taxon_pft_lookup$commonName[i] = "Cascade barberry"}

  else if(genusSpecies == "Mangifera indica"){
    taxon_pft_lookup$pft_MRA[i] = "NA"       # ???   # broadleaf evergreen tree
    taxon_pft_lookup$pft_MODIS[i] = ""
    taxon_pft_lookup$phenology[i] = ""
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = ""
    taxon_pft_lookup$commonName[i] = "mango"}

  else if(genusSpecies == "Melia azedarach"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = ""
    taxon_pft_lookup$phenology[i] = ""
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = ""
    taxon_pft_lookup$commonName[i] = "Chinaberrytree"}

  else if(genusSpecies == "Morus rubra"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = ""
    taxon_pft_lookup$phenology[i] = ""
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = ""
    taxon_pft_lookup$commonName[i] = "red mulberry"}

  else if(genusSpecies == "Nyssa aquatica"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "water tupelo"}

  else if(genusSpecies == "Nyssa biflora"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "swamp tupelo"}

  else if(genusSpecies == "Nyssa sylvatica"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "blackgum"}

  else if(genusSpecies == "Oxydendrum sp."){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "sourwood"}

  else if(genusSpecies == "Picea engelmannii"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "Engelmann spruce"}

  else if(genusSpecies == "Picea glauca"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "white spruce"}

  else if(genusSpecies == "Picea mariana"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "black spruce"}

  else if(genusSpecies == "Picea sp."){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "spruce"}

  else if(genusSpecies == "Pinus contorta"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "lodgepole pine"}

  else if(genusSpecies == "Pinus echinata"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "shortleaf pine"}

  else if(genusSpecies == "Pinus flexilis"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "limber pine"}

  else if(genusSpecies == "Pinus palustris"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "longleaf pine"}

  else if(genusSpecies == "Pinus ponderosa"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "ponderosa pine"}

  else if(genusSpecies == "Pinus sp."){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "pine"}

  else if(genusSpecies == "Pinus strobus"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "eastern white pine"}

  else if(genusSpecies == "Pinus taeda"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "loblolly pine"}

  else if(genusSpecies == "Pisonia albida"){
    taxon_pft_lookup$pft_MRA[i] = "NA"    # Growth Form - Multiple Stem, Height, Mature - 30ft
    taxon_pft_lookup$commonName[i] = "corcho bobo"}

  else if(genusSpecies == "Platanus occidentalis"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "American sycamore"}

  else if(genusSpecies == "Populus deltoides"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "eastern cottonwood"}

  else if(genusSpecies == "Populus grandidentata"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "bigtooth aspen"}

  else if(genusSpecies == "Populus sp."){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "cottonwood"}

  else if(genusSpecies == "Populus tremuloides"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "quaking aspen"}

  else if(genusSpecies == "Prosopis glandulosa"){
    taxon_pft_lookup$pft_MRA[i] = "NA"   # shrub or tree ranging from 4 to 6m tall
    taxon_pft_lookup$commonName[i] = "honey mesquite"}

  else if(genusSpecies == "Prosopis velutina"){
    taxon_pft_lookup$pft_MRA[i] = "NA" # deciduous shrub or tree is less than 15 m
    taxon_pft_lookup$commonName[i] = "velvet mesquite"}

  else if(genusSpecies == "Prunus americana"){
    taxon_pft_lookup$pft_MRA[i] = "NA" # deciduous large shrub or small tree, reaching heights up to 15 feet
    taxon_pft_lookup$commonName[i] = "American plum"}

  else if(genusSpecies == "Prunus angustifolia"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub" # short, thickly branched shrub 3 to 10ft tall
    taxon_pft_lookup$commonName[i] = "Chickasaw plum"}

>>>>

I searched for each taxon on USDA plants and determined the phenology and leaf type. 
- shrubs that are angiosperms are broadleaf
- group = dicot or monocot, you can say broadleaf
- group = gymnosperm will be needleleaf

Utilized the NEON vst_apparentindividual data for growthform.

```{r manually_add_pft_labels}
# list unique taxon labels, add an empty columns to fill with plant functional type (PFT) and common name labels.
taxon_pft_lookup <- data.frame(genusSpecies = unique(cleaned_spectra$genusSpecies),
                               commonName = NA,
                               pft_MRA = NA,
                               pft_MODIS = NA,
                               phenology = NA,
                               leafType = NA,
                               growthForm = NA,
                               comments = "") %>% 
  # arrange alphabetically
  dplyr::arrange(genusSpecies)

@@

  else if(genusSpecies == "Acacia greggii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub" 
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "catclaw acacia"}

@@

  else if(genusSpecies == "Aloysia wrightii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "Wright's beebrush"}

@@

  else if(genusSpecies == "Artemisia tridentata"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "big sagebrush"}

  else if(genusSpecies == "Atriplex canescens"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "fourwing saltbush"}

  else if(genusSpecies == "Atriplex confertifolia"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "shadscale saltbush"}

@@

  else if(genusSpecies == "Bourreria sp."){
    taxon_pft_lookup$pft_MRA[i] = "NA"
    taxon_pft_lookup$pft_MODIS[i] = "NA"
    taxon_pft_lookup$phenology[i] = "NA"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "strongbark"
    taxon_pft_lookup$commonName[i] = "varied genus"}

@@

  else if(genusSpecies == "Chrysothamnus viscidiflorus"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "yellow rabbitbrush"}

@@

  else if(genusSpecies == "Cornus drummondii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub" 
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "roughleaf dogwood"}

@@

  else if(genusSpecies == "Corylus cornuta"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub" 
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "beaked hazelnut"}

@@

  else if(genusSpecies == "Ephedra trifurca"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "longleaf jointfir"}

  else if(genusSpecies == "Ericameria laricifolia"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "turpentine bush"}

@@

  else if(genusSpecies == "Gaultheria shallon"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "salal"}

@@

  else if(genusSpecies == "Gutierrezia sarothrae"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "broom snakeweed"}

@@

  else if(genusSpecies == "Juniperus osteosperma"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "evergreen_needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "Utah juniper"
    taxon_pft_lookup$comments[i] = "scale-like needle leaves"}

  else if(genusSpecies == "Juniperus virginiana"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "evergreen_needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf" 
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "red cedar"
    taxon_pft_lookup$comments[i] = "scale-like needle leaves"}

  else if(genusSpecies == "Larix sp."){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "larch"
    taxon_pft_lookup$comments[i] = "unusual genus with deciduous conifers"}

  else if(genusSpecies == "Larrea tridentata"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "creosote bush"}

  else if(genusSpecies == "Liquidambar styraciflua"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "sweetgum"}

  else if(genusSpecies == "Liriodendron tulipifera"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "tuliptree"}

  else if(genusSpecies == "Lonicera maackii"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "Amur honeysuckle"}

  else if(genusSpecies == "Maclura pomifera"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree" 
    taxon_pft_lookup$commonName[i] = "Osage-orange"}

 else if(genusSpecies == "Magnolia sp."){
    taxon_pft_lookup$pft_MRA[i] = "NA"
    taxon_pft_lookup$pft_MODIS[i] = "NA"
    taxon_pft_lookup$phenology[i] = "NA"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "NA"
    taxon_pft_lookup$commonName[i] = "magnolia"
    taxon_pft_lookup$comments[i] = "varied genus with both deciduous and evergreen trees and shrubs"}

  else if(genusSpecies == "Mahonia nervosa"){
    taxon_pft_lookup$pft_MRA[i] = "woody_shrub"
    taxon_pft_lookup$pft_MODIS[i] = "shrub"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "Cascade barberry"}

  else if(genusSpecies == "Mangifera indica"){
    taxon_pft_lookup$pft_MRA[i] = "NA" # not coniferous or deciduous
    taxon_pft_lookup$pft_MODIS[i] = "evergreen-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "mango"}

  else if(genusSpecies == "Melia azedarach"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "Chinaberrytree"}

  else if(genusSpecies == "Morus rubra"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "red mulberry"}

  else if(genusSpecies == "Nyssa aquatica"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "water tupelo"}

  else if(genusSpecies == "Nyssa biflora"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "swamp tupelo"}

  else if(genusSpecies == "Nyssa sylvatica"){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "blackgum"}

  else if(genusSpecies == "Oxydendrum sp."){
    taxon_pft_lookup$pft_MRA[i] = "deciduous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "deciduous-broadleaf_forest"
    taxon_pft_lookup$phenology[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "sourwood"}

  else if(genusSpecies == "Picea engelmannii"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "evergreen-needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "Engelmann spruce"}

  else if(genusSpecies == "Picea glauca"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "evergreen-needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "white spruce"}

  else if(genusSpecies == "Picea mariana"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "evergreen-needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "black spruce"}

  else if(genusSpecies == "Picea sp."){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "evergreen-needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "spruce"}

  else if(genusSpecies == "Pinus contorta"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "evergreen-needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "lodgepole pine"}

  else if(genusSpecies == "Pinus echinata"){
    taxon_pft_lookup$pft_MRA[i] = "coniferous_forest"
    taxon_pft_lookup$pft_MODIS[i] = "evergreen-needleleaf_forest"
    taxon_pft_lookup$phenology[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "tree"
    taxon_pft_lookup$commonName[i] = "shortleaf pine"}
